% !TEX root=../MA119-Main.tex


\begin{tcolorbox}[colback=white,colframe=cyan, title filled=false, coltitle=cyan, enhanced, attach boxed title to top center={yshift=-3mm,yshifttext=-1mm}, fonttitle=\bfseries, boxed title style={size=small,colback=white},
title=Use logarithms to solve exponential equations
]
\begin{enumerate}
\item  Rewrite the equation in the form $b^u=c$, where $u$ is an algebraic expression. 
\item Take logarithm of both sides. If $b=10$, take common logarithm. Otherwise, take the natural logarithm.
\item Simplify the resulting equation $u=\log_bc$ and solve for the variable. 
\end{enumerate}
\end{tcolorbox}

\begin{tcolorbox}[colback=white,colframe=cyan, title filled=false, coltitle=cyan, enhanced, attach boxed title to top center={yshift=-3mm,yshifttext=-1mm}, fonttitle=\bfseries, boxed title style={size=small,colback=white},
title=Use exponential  form to solve logarithmic equations
]
\begin{enumerate}
\item Rewrite the equation in the form $\log_bu=c$, where $u$ is an algebraic expression. 
\item Using the definition of logarithm to rewrite the equation in the form $u=b^c$.
\item Solve the resulting equation for the variable. 
\item check proposed solutions in the original equation.
\end{enumerate}
\end{tcolorbox}




\begin{exercise}
Solve each exponential equation

\noindent
\begin{enumerate*}[label=(\arabic*)~~]
\item $4^{2x-1}=64$\hspace{0.4\textwidth}
  
\item $2\cdot 10^{x}-1=3$ \quad\quad $x=\log 2$
\end{enumerate*}
\end{exercise}

\begin{exercise}
Solve each logarithmic equation.

\noindent 
\begin{enumerate*}[label=(\arabic*)~~]
\item $\log(2x+50)=2$\hspace{0.3\textwidth}
\item $\log_2 \sqrt{x-1}=1$
\end{enumerate*}

\vspace{2.5cm}
\noindent
\begin{enumerate*}[resume*]
\item $\log_5x+\log_5(4x-1)=1$\hspace{0.26\textwidth}
\item $\log(16-20x)=2+\log(-x)$
\end{enumerate*}
\end{exercise}
\vspace{2cm}

\begin{exercise}
Using the formula $A=P(1+\frac rn)^{nt}$ to determine how many year, to the nearest tenth, it will take \$1000 to \$3600 at the interest rate 8\% compounded quarterly.
\end{exercise}
$t\approx 16.2$

Answer to the practice problem is 13.89. 

